---
title: Points Model
output: 
  html_document:
    toc: false
    toc_depth: 3
    toc_float: true
    code_folding: hide
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
params:
  min_age: 18
  input_table: covid_points_model_population_2021_12_25__2022_03_16_no_pax

  outcome_col_name: first_time_severe
  create_tables: 1
  plot_hists: 0
  sanity_tests: 1
  fwp_window: 21
  calc_broom_fit_ci: 1
  filter_missing_bmi: 0
  seed: 2013
  debug_mode: 1
  train_prop: 0.8
  remove_paxlovid_takers: 1

---

## parameters:
  * input_table =  `r params$input_table`\
  * min_age =  `r params$min_age`\
  * outcome columns =  `r params$outcome_col_name`\
  * outcome = `r ifelse (params$outcome_col_name == 'first_time_severe' , 'Severe or Death', 'Hospitalizaiton ? ')` 
  * fwp_window = `r params$fwp_window` days \
  * calc_broom_fit_ci = `r params$calc_broom_fit_ci`\
  * train_prop = `r params$train_prop`\ 
  * filter_missing_bmi: `r params$filter_missing_bmi`\
  * remove_paxlovid_takers: `r params$remove_paxlovid_takers`\

```{r message=FALSE, warning=FALSE, include=FALSE}
# 
 library(plyr)
 library(tidyverse)
 library(RODBC)
 library(mgcv)
 library(gtsummary)
 library(ggplot2)
 library(MatchIt)
 library(Hmisc)
 library(readxl)
 library(glue)
 library(dplyr)
 library(feather)
 library(cobalt)
 library(knitr)
 library(cri.utils)
 library(lubridate)
 library(jtools)
 library(flextable)
 library(formula.tools)
 library(fastDummies) # to create dummy columns
 library(pROC)
 source('functions.R')
 library(dbplyr)
```


```{r include=FALSE}
options(scipen = 999)
activate_keytab("innovationdev1")
set.seed(params$seed)

```

```{r include=FALSE}
con <- cri_connection()


```


```{r read sql,  include=FALSE}
filename = glue("feathers/{params$input_table}.feather")
if (file.exists(filename)) {
  df_raw = read_feather (filename)
} else {
df_raw = connect_table(con, "Mechkar.[clalit\\yatirbe]", params$input_table) %>%
  select(-teudat_zehut, -t_index_date) %>%
  mutate(index_date = positive_date) %>% 
  rename(date_ptria_corona = covid_death_date) %>% 
  rename(sw_previous_infection = sw_reinfection) %>% 
  distinct() %>%
  collect()
write_feather(df_raw, filename)
}
```



First PCR Positive date: `r min(df_raw$index_date) `
Last PCR Positive date: `r max(df_raw$index_date) `



```{r eval=FALSE, include=FALSE}
describe(df_raw)
```


```{r initial cleaning, echo=FALSE}
df_raw = df_raw %>%
  dplyr::mutate(psw_nursing_home = replace_na(psw_nursing_home, 0)) %>%
  dplyr::mutate(age = pn_age) %>%
  dplyr::mutate(sex = factor(sex)) %>% 
  dplyr::mutate(sex = fct_relevel(sex, 'F','M')) %>%
  dplyr::mutate(days_from_dose_1 =  ifelse(days_from_dose_1 < 0, as.numeric(NA_integer_),days_from_dose_1)) %>% 
  print_total("")
  

```

```{r sanity on dose1 date, echo=FALSE}
num_invalid_dose1_date =  df_raw %>%
  filter(days_from_dose_1 < 0) %>% 
  select(days_from_dose_1) %>%
  dplyr::summarise(n=dplyr::n()) %>% 
  pull(n)

stopifnot(params$sanity_tests == 0 || num_invalid_dose1_date ==0)
```


```{r age and membership filtering, echo=FALSE}
  df = df_raw %>%
  print_total("before age filtering") %>% 
    filter(between(age, params$min_age, 150)) %>% 
  print_total("after age filtering") %>% 
  filter(sw_cont_membership == 1) %>%
  print_total("after cont. membership filtering") 

```
```{r paxlovid filtering, echo=FALSE}  
if (params$remove_paxlovid_takers == 1) {
  df = df %>% 
  filter(pax_taker == 0 ) %>%
  print_total("after filtering Paxlovid takers")
} else {
  print("Avoding the removal of paxlovid takes")
}
```

```{r echo=FALSE}
if (params$debug_mode == 1){
  
  df %>%
    filter(first_time_severe < index_date) %>%
    select(index_date, first_time_severe )
}
```



```{r time to outcome calc, echo=FALSE}
fwp_end_date =  max(as.Date((df_raw$retrieval_date)))
df = df %>%
 mutate(days_to_end_of_fwp = as.numeric(difftime(fwp_end_date, index_date, units = "days"))) 
 
```

```{r time to outcome , echo=FALSE}
if (params$plot_hists == 1) {
hist(df$days_to_end_of_fwp, breaks=40, freq=F)
}
```

```{r}
if (params$plot_hists == 1) {
  tmp = df %>%
    filter(pn_count_hosp>3)
hist(tmp$pn_count_hosp, breaks=30, freq=T,xlim =c(0,20), main = "hist of pn_count_hosp when > 3")
}


```
```{r}
if (params$plot_hists == 1) {
  N = nrow(df)
 df %>% 
    group_by(pn_count_hosp) %>% 
    count() %>% 
    arrange(pn_count_hosp) %>% 
    mutate(percent = 100*round(cumsum(n)/N,3)) %>% 
    kable()
    
  
}
```


```{r}
outcome_col = params$outcome_col_name

### determining the outcome accodring to the followup window
fwp_window = params$fwp_window
df = df %>% 
 print_total("before asserting a minimal fwp window ") %>%
 filter(days_to_end_of_fwp >= fwp_window) %>% 
 print_total("after asserting a minimal fwp window ") %>%
 mutate (days_to_outcome    = as.numeric(difftime(!!sym(outcome_col), index_date,       units = "days"))) %>%
 filter (is.na(days_to_outcome) | days_to_outcome >= 0) %>%
 print_total("after filtering outcome events occuring before index_date ") %>% 
 mutate(sw_outcome = case_when(between(days_to_outcome, 0 ,  fwp_window) ~ 1, T ~0))
```


```{r}
hist(df$days_to_outcome)
```

```{r}
df %>% 
  filter(is.na(bmi)) %>% 
  count() %>% 
  print_total("entries with missing bmi")  
```




### outcome events
```{r}
table(df$sw_outcome)
```
```{r loess curve for age, warning=FALSE}
if (params$plot_hists == 1) {
  span = 0.2
  tmp = df %>% 
    slice_sample(n=55000)
  plot(tmp$age  ,predict(loess(sw_outcome ~ age,tmp, span=span)),ylab='P(severe covid19)', xlab='Age (years)',
       main = sprintf('loess curves with span=%.1f', span))

}
```



```{r loess curve for count_hosp, warning=FALSE}
if (params$plot_hists == 1) {
  span = 0.4
  tmp = df %>% 
    slice_sample(n=100000)
  plot(tmp$pn_count_hosp,predict(loess(sw_outcome ~ pn_count_hosp,tmp, span=span)),ylab='P(severe covid19)', xlab='pn_count_hosp',
       main = sprintf('loess curves with span=%.1f', span))
}
```
```{r}
colnames(df)
max(df$first_time_severe, na.rm = T)
```


```{r eval=FALSE, include=FALSE}

df %>% 
  group_by(days_to_outcome) %>% 
  count()
```

```{r months from variables}
df = df %>%
  mutate(days_from_last_vac_dose = pmin(days_from_dose_2,days_from_dose_3,days_from_dose_4, na.rm =T)) %>% 
  mutate(months_from_last_vac_dose =  ceil(days_from_last_vac_dose/30.0)) %>% 
  mutate (months_from_last_vac_dose_cat = case_when(  
                              between(months_from_last_vac_dose, 1, 6)  ~ as.character('1_6'),
                              between(months_from_last_vac_dose, 7, 10)  ~ as.character('7_10'),
                                  months_from_last_vac_dose >= 11 ~ as.character('11plus') ,
                              T ~  as.character('unvacc'))) %>% 
  mutate (months_from_last_vac_dose_cat = fct_relevel(months_from_last_vac_dose_cat, 'unvacc','1_6', '7_10', '11plus')) %>%
  mutate(months_from_2nd_dose = round(days_from_dose_2/30.0)) %>% 
  mutate(months_from_2nd_dose_cat = case_when(
                                    between(months_from_2nd_dose, 0, 3) ~ as.character('0_3'),
                                    between(months_from_2nd_dose, 4, 5) ~ as.character('4_5'),
                                    months_from_2nd_dose >= 6 ~ as.character('6plus'),
                                    T ~ as.character('unvacc'))) %>% 
  mutate (months_from_2nd_dose_cat = fct_relevel(months_from_2nd_dose_cat, 'unvacc','0_3','4_5','6plus')) %>%
  # Note as of this modeling time (Jan 2022) there are not enough people with more than 6 months from 3rd dose
  mutate(months_from_3rd_dose = round(days_from_dose_3/30.0)) %>% 
  mutate(months_from_3rd_dose_cat = case_when(
                                    between(months_from_3rd_dose, 0, 1) ~ as.character('0_1'),
                                    between(months_from_3rd_dose, 2, 3) ~ as.character('2_3'),
                                    months_from_3rd_dose >= 4 ~ as.character('4plus'),
                                    T ~ as.character('unvacc'))) %>% 
  mutate (months_from_3rd_dose_cat = fct_relevel(months_from_3rd_dose_cat, 'unvacc','0_1', '2_3','4plus')) %>%
  
  mutate(days_from_prev_infection = as.numeric(difftime(index_date, previous_infection_date, units = "days"))) %>% 
  mutate(months_from_prev_infection = round(days_from_prev_infection/30.0)) %>% 
  mutate (months_from_prev_infection_cat = case_when(  
            between(months_from_prev_infection, 1, 5) ~ as.character('1_5'),
            between(months_from_prev_infection, 6, 8) ~ as.character('6_8'),
            between(months_from_prev_infection, 9, 11) ~ as.character('9_11'),
            between(months_from_prev_infection, 12, 14) ~ as.character('12_14'),
            months_from_prev_infection >= 15 ~ as.character('15plus') ,
            T ~  as.character('None'))) %>% 
  mutate (months_from_prev_infection_cat = fct_relevel(months_from_prev_infection_cat, 'None','1_5', '6_8','9_11','12_14','15plus')) %>% 
  mutate(days_from_last_immune_challenge = pmin(days_from_dose_2, days_from_dose_3,days_from_dose_4, days_from_prev_infection, na.rm = T)) %>% 
  mutate(months_from_last_immune_challenge =  ceil(days_from_last_immune_challenge/30.0)) %>% 
  mutate (months_from_last_immune_challenge_cat = case_when(  
                              between(months_from_last_immune_challenge, 1, 4) ~ as.character('1_4'),
                              between(months_from_last_immune_challenge, 5, 6) ~ as.character('5_6'),
                                  months_from_last_immune_challenge >= 7 ~ as.character('7plus') ,
                              T ~  as.character('unvacc'))) %>% 
  mutate (months_from_last_immune_challenge_cat = fct_relevel(months_from_last_immune_challenge_cat, 'unvacc','1_4', '5_6', '7plus')) %>%
  mutate(days_from_last_dose = pmin(days_from_dose_2, days_from_dose_3, na.rm = T)) %>% 
  mutate(waned_immunity2 =  case_when((num_vacc_doses >=2 | sw_previous_infection == 1) & between(days_from_last_immune_challenge , 7, 150) ~ 0, T ~ 1)) %>%
  mutate(waned_immunity = case_when(num_vacc_doses >=2  & between(days_from_last_dose , 7, 150) ~ 0,                                                                      T                                      ~ 1)) %>% 
  mutate(recent_immunity = 1-waned_immunity2)
```


```{r eval=FALSE, include=FALSE}
manual = df %>%
  select(contains('date'), contains('immune_challenge'), contains('doses'), -contains('corona'), -retrieval_date)

describe(manual)
```


```{r}
df = df %>% 
 mutate(age_group = case_when(
    between(age, 0, 17) ~  "0_17",
    between(age, 18, 29) ~ "18_29",
    between(age, 30, 39) ~ "30_39",
    between(age, 40, 44) ~ "40_44",
    between(age, 45, 49) ~ "45_49",
    between(age, 50, 54) ~ "50_54",
    between(age, 55, 59) ~ "55_59",
    between(age, 60, 64) ~ "60_64",
    between(age, 65, 69) ~ "65_69",
    between(age, 70, 74) ~ "70_74",
    between(age, 75, 79) ~ "75_79",
    between(age, 80, 84) ~ "80_84",
    between(age, 85, 89) ~ "85_89",
    age >= 90 ~ "90plus")) %>% 
mutate(age_group10 = case_when(
    between(age, 0, 17) ~  "bug",
    between(age, 18, 29) ~ "18_29",
    between(age, 30, 39) ~ "30_39",
    between(age, 40, 49) ~ "40_49",
    between(age, 50, 59) ~ "50_59",
    between(age, 60, 69) ~ "60_69",
    between(age, 70, 79) ~ "70_79",
    age >= 80 ~ "80plus")) %>% 
  mutate(num_vacc_doses_numeric = num_vacc_doses) %>% 
  mutate(num_vacc_doses = as.factor(num_vacc_doses)) %>% 
  mutate (num_vacc_doses = fct_relevel(num_vacc_doses, '0','1', '2','3')) %>%  # 4th dose will also exist as last
  mutate(vaccination_status = case_when(num_vacc_doses_numeric <2 ~ as.character(num_vacc_doses_numeric),
                                        num_vacc_doses_numeric >=2 ~ '2plus')) %>% 
  mutate (vaccination_status =fct_relevel(vaccination_status, '0','1', '2plus')) %>%  
 mutate(bmi_30plus = case_when(bmi >= 30 ~ 1, T ~ 0)) # Note missing BMI values will be implicitly imputed to 0

  
```


```{r}
df = df %>% 
 mutate(count_hosp_cat = case_when(
    between(pn_count_hosp, 0, 0) ~  "0",
    between(pn_count_hosp, 1,2) ~  "1_2",
    between(pn_count_hosp, 3, 5) ~ "3_5",
    between(pn_count_hosp, 6, 99999) ~ "6plus"))

  
```





```{r missing values bmi, eval=FALSE, include=FALSE}
void = df %>% 
  print_total( "WITH entries with missing BMI") %>% 
  filter(!is.na(bmi)) %>%
  print_total( "WITHOUT entries with missing BMI") 
```
### report smoking Missing Values Stats
```{r missing values smoking}  
void = df %>% 
  print_total( "WITH entries with smoking status") %>% 
  filter(!is.na(pc_smoke_desc)) %>% 
  print_total( "WITHOUT entries with missing smoking status")
```




```{r applying complete cases on BMI and smoking, eval=FALSE, include=FALSE}
if (params$filter_missing_bmi ==1) {
df = df %>% 
  filter(!is.na(bmi)) %>%
  print_total( "after removing entries with missing BMI") 
} else {
    print_total( "entires with missing BMI is not filtered. (for obesity, will be assumed as 0 )") 
  df = df
}
```


### outcome events after filtering missing BMI
```{r}
table(df$sw_outcome)
```

```{r eval=FALSE, include=FALSE}
df %>% 
  select(age_group,htn,sw_outcome) %>% 
  group_by(age_group,htn) %>% 
  summarise(count=n(),risk =mean(sw_outcome))
```

```{r}
df = df %>% 
  mutate(respiratory = case_when(asthma == 1 | other_resp == 1 ~ 1, T ~ 0)) %>% 
  mutate(age_over_40 = case_when(age<40 ~0, T ~ age-40))
```
```{r eval=FALSE, include=FALSE}
tmp = df %>% 
  select(sw_previous_infection, sex,age,heart,,preg,overweight,obese,severe_obese,
                    sickle, thalasamia, diabetes, respiratory, liver, copd, immunosuppression, waned_immunity, num_vacc_doses_numeric, sw_outcome)
    
Hmisc::rcorr(as.matrix(tmp), type="pearson")
  
```
```{r eval=FALSE, include=FALSE}
describe(df)
```





# Table1
#### Initial population stratified by outcome
```{r table1, echo=FALSE}
if (params$create_tables==1) {
  table1 = tbl_summary(
      df %>% select(age_group10, sex, bmi_30plus,active_cancer, ckd, diabetes,  heart,neuro,liver, copd, immunosuppression,  count_hosp_cat,  months_from_last_vac_dose_cat, sw_previous_infection, sw_outcome),
      by = 'sw_outcome',
      label = list(age_group10 ~ "Age group",
                 sex ~ "Sex",   
                 bmi_30plus ~ "Obesity",
                 active_cancer ~"Active cancer",
                 heart ~ "Heart disease",
                 diabetes ~ "Diabetes",
                 neuro ~ "Neurological disease",
                 liver ~ "Liver disease",
                 count_hosp_cat ~ "3 prior years hospitalization count",
                 months_from_last_vac_dose_cat ~ "Months from last vaccination dose",
                 sw_previous_infection ~ "Previous infection"
                 ),
      statistic = all_continuous() ~ "{mean} ({sd}), {median} ({p25},{p75})",
      missing = "ifany", # don't list missing data separately
    ) %>%
      modify_header(label = "**Variable**") %>% # update the column header
      bold_labels() %>% 
      modify_header(all_stat_cols() ~ "**{ifelse(level==0,'Non Severe Outcome','Severe Outcome')}**, N = {n} ({style_percent(p, symbol = TRUE)})") %>% 
      add_overall() 

table1 %>% 
    gtsummary::as_flex_table() %>% 
    save_as_docx(path= glue("points_model_table1.docx"))
table1
}


```


# fit model


```{r}
df = df %>% 
    mutate(rowid = row_number()) %>% 
  dummy_cols(select_columns = c( 'sex','num_vacc_doses','vaccination_status',  'count_hosp_cat','months_from_last_immune_challenge_cat','months_from_last_vac_dose_cat', 'months_from_prev_infection_cat', 'age_group','age_group10'), remove_first_dummy = TRUE) %>% 
  mutate(age_div_2 = round(age/2.0)) 

data_train = df %>% 
        slice_sample(prop = params$train_prop)

data_test = df %>% anti_join(data_train, by='rowid')
```

As evidence for the determinism of the slice_sample, using a set.seed():
```{r}
sum(data_train$rowid)
```

```{r}
write_feather(data_train, glue("feathers/data_train.feather"))
write_feather(data_test, glue("feathers/data_test.feather"))
```



```{r message=FALSE, warning=FALSE}
predictors = c("age_group10_30_39","age_group10_40_49",
"age_group10_50_59",
"age_group10_60_69",
"age_group10_70_79",
"age_group10_80plus","bmi_30plus", "active_cancer","ckd","diabetes", "heart", "neuro","liver","copd", "immunosuppression",
"count_hosp_cat_1_2","count_hosp_cat_3_5",  "count_hosp_cat_6plus")

formula = reformulate(termlabels = predictors,intercept = TRUE, response = 'sw_outcome' )
Reduce(paste, deparse(formula))


```
# Missing values status
```{r}
  df %>%
  select(age_group10, bmi_30plus,active_cancer, ckd, diabetes,  heart,neuro,liver, copd, immunosuppression,  count_hosp_cat,  months_from_last_vac_dose_cat, sw_previous_infection, sw_outcome) %>% 
  summarize_all(funs(sum(is.na(.)))) %>% 
  pivot_longer(everything(), names_to = "var", values_to = "count") %>% 
  filter(count>0) %>% 
  arrange(desc(count))
```

```{r manual interactions, eval=FALSE, include=FALSE}
formula_interaction = sw_outcome ~ age_group_19_29 + age_group_30_39 + age_group_40_49 +      age_group_50_59 + age_group_60_69 + age_group_70_79 + age_group_80plus +      sw_previous_infection + active_cancer + sex_M + ckd + overweight +      obese + severe_obese + preg + thalasamia + diabetes + respiratory +      htn + neuro + liver + heart + copd + immunosuppression +      pn_count_hosp + months_from_last_immune_challenge_cat * num_vacc_doses

```

```{r eval=FALSE, include=FALSE}
data_test %>% count()
describe(data_train %>% select(all_of(predictors)))
```

# Colineraity check via alias()
```{r eval=FALSE, include=FALSE}
 res = alias(glm(formula ,data = data_train), partial=T)
 res
```

# VIF
```{r}
if (length(predictors) > 2) 
  car::vif(lm(formula ,data = data_train)) %>% kable()
```
# GLM

predictors = `r predictors`
```{r glm, echo=FALSE, message=FALSE, warning=FALSE}
glm_fit = glm(formula, data = data_train, family = "binomial", control=glm.control(maxit=50))
```


```{r include=FALSE}
save(glm_fit, file = "glm_fit.rda")
write_feather(df, glue("feathers/final_df_{params$input_table}.feather"))
```

```{r echo=FALSE}
 summary(glm_fit)
```

```{r broom_tidy, message=FALSE, warning=FALSE}
if (params$calc_broom_fit_ci ==1) {
  tidy_fit = broom::tidy(glm_fit, conf.int=T) %>% 
    print_total("after broom tidy") %>% 
    dplyr::mutate(OR  = estimate, `OR 95% Low` = conf.low, `OR 95% High` = conf.high ) %>% 
    dplyr::mutate(OR  = estimate, `estimate 95% Low` = conf.low, `estimate 95% High` = conf.high ) %>% 
    dplyr::mutate(across(c("OR","OR 95% Low","OR 95% High"), exp)) %>% 
    dplyr::mutate(across(c("OR","OR 95% Low","OR 95% High"),  ~round(.x, 3))) %>% 
    dplyr::mutate(across(c("estimate","estimate 95% Low","estimate 95% High"),  ~round(.x, 3))) %>% 
    dplyr::mutate(across(c("p.value"), ~round(.x, 5))) %>% 
  select(term, contains('estimate'),p.value, contains('OR') )

} else {
  tidy_fit = broom::tidy(glm_fit, conf.int=F) %>% 
    dplyr::mutate(exp_estimate = exp(estimate)) %>% 
    dplyr::mutate(across(c("estimate", "exp_estimate"), ~round(.x, 2))) %>% 
    dplyr::mutate(p.value = round(p.value,5)) %>% 
  select(term, estimate, exp_estimate,p.value)  
  
}
write.csv(tidy_fit, "logistic_regression_model.csv")

tidy_fit %>% kable()
```


